Phoneme surprisal

### 2019-06-21 surprisal/phoneme_surprisal.Rmd
hold out entropy

```{r setup, include=FALSE}
library(tidytext)
library(tidyverse)
library(childesr)
library(glue)
library(feather)
library(here)

@@

adult_phonemes <- adult_utterances %>%
  left_join(phones) %>%
  filter(!is.na(ipa)) %>%
  unnest_tokens(phone, ipa, token = stringr::str_split, 
                pattern = "_+", drop = F) %>%
  group_by(utterance_id, word_id) %>%
  mutate(phone_order = 1:n()) %>%
  mutate(word_length = max(phone_order)) %>%
  ungroup() %>%
  select(utterance_id, word_id, word, ipa, phone, phone_order, word_length)

adult_phonemes %>%
  filter(phone_order == 1) %>%
  ggplot(aes(x = word_length)) + 
  geom_histogram(fill = "white", color = "black") + 
  theme_classic()

@@

```{r}
four_lengths <- adult_phonemes %>%
  filter(word_length == 3) %>%
  group_by(utterance_id, word_id, phone_order)

four_words <- four_lengths %>%
  ungroup() %>%
  distinct(word, phone_order, phone) %>%
  group_by(word) %>%
  nest()

word1 <- four_words %>%
  slice(1) %>%
  unnest() %>%
  slice(c(2,3)) %>%
  select(-word)

word1 %>%
  spread(phone_order, phone) %>%
  left_join(four_lengths %>%
  spread(phone_order, phone)) %>%
  distinct(utterance_id, word_id, word) %>%
  group_by(word) %>%
  summarise(n = n())

>>>>

```{r setup, include=FALSE}
library(tidytext)
library(tidyverse)
library(childesr)
library(glue)
library(feather)
library(here)
library(entropy)
library(tidyboot)
library(janitor)
library(arrangements)

@@

adult_phonemes <- adult_utterances %>%
  left_join(phones) %>%
  filter(!is.na(ipa)) %>%
  unnest_tokens(phone, ipa, token = stringr::str_split, 
                pattern = "_+", drop = F) %>%
  filter(phone != "") %>%
  group_by(utterance_id, word_id) %>%
  mutate(phone_order = 1:n()) %>%
  mutate(word_length = max(phone_order)) %>%
  ungroup() %>%
  select(utterance_id, word_id, word, ipa, phone, phone_order, word_length)

adult_phonemes %>%
  filter(phone_order == 1) %>%
  ggplot(aes(x = word_length)) + 
  geom_histogram(fill = "white", color = "black", binwidth = 1) + 
  theme_classic()

@@

```{r leave-out-surprisal}
hold_out_entropy <- function(df) {
  spread_df <- df %>%
    spread(phone_order, phone) %>%
    clean_names()

  phone_length <- max(df$phone_order)

  lens <- paste0("x", 1:phone_length)

  boot_entropy <- function(group) {
    spread_df %>%
      group_by_at(group) %>%
      summarise(n = n()) %>%
      summarise(entropy = entropy(n)) %>%
      ungroup() %>%
      tidyboot_mean(entropy)
  }

  groupings <- combinations(lens, phone_length -1, 
                            layout = "list")

  groupings <- map(seq(length(lens),1,-1),
                   function(x) c(groupings[[x]], lens[x]))

  map_df(groupings, boot_entropy) %>%
    mutate(phone_order = seq(phone_length, 1, -1))

}

split_lengths <- adult_phonemes %>%
  group_by(word_length) %>%
  nest() %>%
  filter(word_length > 1) %>%
  filter(word_length %in% 1:6 ) %>%
  mutate(entropy = map(data, hold_out_entropy))

split_lengths %>%
  select(-data) %>%
  unnest() %>%
  ggplot(aes(x = phone_order, y = empirical_stat)) + 
  facet_wrap(~ word_length, scales = "free") +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  theme_classic() + 
  geom_line()
```
